---
title: "Using fintrackr for Personal Finance Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using fintrackr for Personal Finance Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(fintrackr)
library(ggplot2)
```


## Introduction

The fintrackr package provides tools for personal finance management,
including budgeting, investment growth analysis, debt repayment projection, 
and financial visualization.

## Example Data


```{r}
finance_data <- data.frame(
  date = as.Date(c(
    "2026-01-01","2026-01-05","2026-01-10","2026-01-15","2026-01-20",
    "2026-02-01","2026-02-03","2026-02-08","2026-02-12","2026-02-18",
    "2026-03-01","2026-03-06","2026-03-10","2026-03-15","2026-03-22"
  )),
  category = c(
    "Salary","Food","Transport","Groceries","Entertainment",
    "Salary","Rent","Food","Transport","Shopping",
    "Salary","Groceries","Food","Utilities","Entertainment"
  ),
  amount = c(
    3000, -50, -30, -200, -100,
    3000, -1200, -60, -40, -150,
    3000, -250, -70, -180, -120
  )
)

finance_data
```

Each row represents a single financial transaction. The amount column 
stores both income and expenses: positive values represent income, 
while negative values represent expenses. This unified format simplifies 
analysis and reflects how financial data is commonly recorded in 
real-world applications.

Categories may appear multiple times (e.g., "Salary") because
transactions occur on different dates, allowing the data to track
financial activity over time.


### Budget Summary

```{r}
# finance_data is passed as 'data' argument
# Setting opening_balance = 0 assumes the user starts from zero,
# and all income and expenses update the balance over time.
budget_out <- budget_summary(finance_data, amount, category, date, opening_balance = 0)

head(budget_out$ledger)
tail(budget_out$ledger)
budget_out$category_summary
budget_out$totals
budget_out$monthly_summary
```
This function summarizes financial activity using a single amount column, where positive values represent income and negative values represent expenses. It calculates category totals, overall totals, and monthly trends, while also maintaining a running balance from an opening balance of 0, helping users clearly track how their financial position changes over time.


### Investment Growth

```{r}
invest_out <- calc_returns(
  principal = 0,
  rate = 0.05,
  time = 5,
  compound_freq = 12,
  contribution = 200
)

invest_out$future_value
head(invest_out$growth_table)
```
This function models investment growth with regular contributions (SIP). At each period, the investment earns interest and a fixed contribution is added, making it useful for simulating real-world long-term savings plans.


### Debt Projection

```{r}
debt_out <- debt_projection(
  balance = 5000,
  annual_rate = 0.18,
  monthly_payment = 200
)

head(debt_out)
tail(debt_out)
```
This function creates an amortization-style debt repayment table and shows
how the balance decreases over time until payoff.



### Financial Trends Plot

```{r}
plot_financial_trends(
  budget_out$monthly_summary,
  month,
  net
)
```


This function returns a ggplot object to visualize financial trends over time. It uses the monthly_summary output from budget_summary to show how net balance changes across months, helping users identify patterns in income and spending.


## Conclusion

The fintrackr package provides a simple and effective way to
analyze personal finances by combining data summarization, financial
modeling, and visualization into a single workflow.
