---
title: "Using fintrackr for Personal Finance Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using fintrackr for Personal Finance Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(devtools)
load_all()
library(ggplot2)
library(fintrackr)
library(ggplot2)
```

## Introduction

The fintrackr package provides tools for personal finance management,
including budgeting, investment growth analysis, debt repayment projection,
strategy comparison, and financial visualization.

## Example Data

```{r}
finance_data <- data.frame(
  date = as.Date(c("2026-01-01", "2026-01-15", "2026-02-01", "2026-02-10")),
  category = c("Salary", "Groceries", "Salary", "Rent"),
  income = c(3000, 0, 3000, 0),
  expense = c(0, 200, 0, 1200)
)

finance_data
```
### Budget Summary

```{r}
budget_out <- budget_summary(finance_data, income, expense, category, date)

budget_out$category_summary
budget_out$totals
budget_out$monthly_summary
```
This function helps users summarize their financial activity by category
and by month, while also computing total income, total expenses, and
net balance.


### Investment Growth

```{r}
invest_out <- calc_returns(
  principal = 1000,
  rate = 0.05,
  time = 5,
  compound_freq = 12
)

invest_out$future_value
head(invest_out$growth_table)
```
This function computes compound investment growth over time and returns
both the final future value and a table of values by period.


### Debt Projection

```{r}
debt_out <- debt_projection(
  balance = 5000,
  annual_rate = 0.18,
  monthly_payment = 200
)

head(debt_out)
tail(debt_out)
```
This function creates an amortization-style debt repayment table and shows
how the balance decreases over time until payoff.

### Strategy Comparison

```{r}
strategies <- data.frame(
  strategy_name = c("Plan A", "Plan B"),
  initial = c(1000, 1000),
  rate = c(0.05, 0.07),
  compound_freq = c(12, 12),
  time = c(5, 5)
)

compare_strategies(strategies)
```


This function compares different financial strategies and returns
their final values for easy comparison.



### Financial Trends Plot

```{r}
plot_financial_trends(
  budget_out$monthly_summary,
  month,
  net
)
```


This plotting function returns a ggplot object, allowing users to
visualize financial trends and customize the plot further if needed.


## Conclusion

The fintrackr package provides a simple and effective way to
analyze personal finances by combining data summarization, financial
modeling, and visualization into a single workflow.
